---
title: "Peer Assessment 1 - Reproducible Research"
author: "Gary W. Deinert"
date: "Saturday, April 18, 2015"
output: html_document
---

### Part 1.  Loading and preprocessing the data.  

```{r}
setwd("//GARY-LAPTOP/Users/Gary/Desktop/Data Science Specialzation/Repr Research Peer Assignmt 1")
my.dataframe <- read.csv("activity.csv",header = TRUE)

```
### Part 2.  What is the mean total number of steps taken per day? 

```{r}
      dayTotal <- aggregate(my.dataframe$steps, list(Date = my.dataframe$date), sum)
      dayMean <- format(mean(dayTotal$x, na.rm = TRUE), nsmall = 2)
      dayMedian <- median(dayTotal$x, na.rm = TRUE)

hist(dayTotal$x, main = "Histogram of Daily Walking Totals", xlab = "Daily Steps per Day")

```

***  __The average number of steps per day is `r dayMean` and the median number of steps per day is `r dayMedian`.__   
  

### Part 3.  What is the average daily activity pattern?  

```{r}
myAvgs <- aggregate(my.dataframe$steps, 
            list(Interval = my.dataframe$interval), 
            mean, 
            na.rm = TRUE)
maxInt <- subset(myAvgs,x == max(myAvgs$x))            ##  Find the maximum average time interval 

plot(myAvgs$Interval, myAvgs$x, 
      type = "l",
      main = "Number of Steps per Daily Time Interval", 
      xlab = "Time Interval During Day",
      ylab = "Steps")

```
   
*** __The maximum average daily steps was in time interval `r maxInt$Interval` with `r format(maxInt$x, digits = 2, nsmall = 2)` average steps.__   
