---
title: "Peer Assessment 1 - Reproducible Research"
author: "Gary W. Deinert"
date: "Saturday, April 18, 2015"
output: html_document
---

### Part 1.  Loading and preprocessing the data.  

```{r Load and Preprocess}
setwd("//GARY-LAPTOP/Users/Gary/Desktop/Data Science Specialzation/Repr Research Peer Assignmt 1")
my.dataframe <- read.csv("activity.csv",header = TRUE)

```
### Part 2.  What is the mean total number of steps taken per day? 

```{r Mean Daily Totals}
      dayTotal <- aggregate(my.dataframe$steps, list(Date = my.dataframe$date), sum)
      dayMean <- format(mean(dayTotal$x, na.rm = TRUE), digits=2, nsmall = 2)
      dayMedian <- median(dayTotal$x, na.rm = TRUE)

hist(dayTotal$x, 
     main = "Histogram of Daily Walking Totals", 
     xlab = "Daily Steps",
     col = "blue")

```

***  __The average number of steps per day is `r dayMean` and the median number of steps per day is `r dayMedian`.__   
  

### Part 3.  What is the average daily activity pattern?  

```{r Averages Thru The Day}
myAvgs <- round(aggregate(my.dataframe$steps, 
                  list(Interval = my.dataframe$interval), 
                  mean, 
                  na.rm = TRUE), 0)

maxInt <- subset(myAvgs,x == max(myAvgs$x))            ##  Find the maximum average time interval 

plot(myAvgs$Interval, myAvgs$x, 
      type = "l",
      main = "Number of Steps per Daily Time Interval", 
      xlab = "Time Interval During Day",
      ylab = "Steps")

```
   
*** __The maximum average daily steps was in time interval `r maxInt$Interval` with `r round(maxInt$x, 0)` average steps.__  
   
### Part 4. Imputing missing values.

``` {r Imputing NAs}

sum(is.na(my.dataframe$steps))      ## count the number of NA's
new.dataframe <- my.dataframe       ## create the new dataset equal to the old, to fill in
                                    ## replace the NA's with overall timeinterval mean
suppressWarnings(
new.dataframe$steps[is.na(new.dataframe$steps)] <- myAvgs$x[myAvgs$Interval == new.dataframe$interval])

      newTotal <- aggregate(new.dataframe$steps, list(Date = new.dataframe$date), sum)
      newMean <- format(mean(newTotal$x, na.rm = TRUE), digits=2, nsmall = 2)
      newMedian <- format(median(newTotal$x, na.rm = TRUE), digits = 1, nsmall = 0)

hist(newTotal$x, 
     main = "Updated Histogram of Daily Walking Totals", 
     xlab = "Daily Steps, NA Imputed to Mean",
     col = "green")
```
   
***  __The average number of steps per day is `r newMean` and the median number of steps per day is `r newMedian`.  Compare to the prior average steps per day at `r dayMean` and previous median of `r dayMedian`.   We conclude that imputing values by the interval mean for this dataset does not significantly change the results.__   

### Part 5.  Are there differences in activity patterns between weekdays and weekends?

```{r Weekends vs Weekdays}

weekly.dataframe <- cbind(new.dataframe, partofweek = "Weekday")  ## default all days to weekday
weekly.dataframe <- cbind(weekly.dataframe, day = weekdays(as.Date(weekly.dataframe$date),abbreviate = TRUE))

weekly.dataframe$day <- gsub("S..","ddd",weekly.dataframe$day)    ##  Change weekend days to "ddd"
weekly.dataframe$partofweek <- as.character(weekly.dataframe$partofweek)  ## r requires conversion
x <- which(weekly.dataframe$day == 'ddd')                                 ## find weekend days
weekly.dataframe$partofweek[x] <- "Weekend"                               ## from factor to char
weekly.dataframe$partofweek <- as.factor(weekly.dataframe$partofweek)     ## then convert back
weekly.dataframe$day <- NULL       ## strip weekday variable

##  find weekday averages and weekend averages
weekdays <- weekly.dataframe[weekly.dataframe$partofweek == "Weekday",]
weekdayAvgs <- round(aggregate(weekdays$steps, 
                  list(interval = weekdays$interval), 
                  mean), 0)

weekends <- weekly.dataframe[weekly.dataframe$partofweek == "Weekend",]
weekendAvgs <- round(aggregate(weekends$steps,
                  list(interval = weekends$interval),
                  mean), 0)

par(mfrow = c(2,1))

plot(weekdays$interval,
     weekdays$steps, 
     type="l",
     main = "Weekdays",
     xlab="Time Interval",
     ylab="Steps")

plot(weekends$interval,
     weekends$steps, 
     type="l",
     main = "Weekends",
     xlab="Time Interval",
     ylab="Steps")
